---
title: "Goreference landing sites for fishery logs and obtain temperature data for each catch point"
output: html_notebook
---


```{r load package, include=FALSE}

# FIRST MOUNT DRIVE
#for error transport endpoint is not connected bucket 
# https://stackoverflow.com/questions/14010115/amazon-s3-with-s3fs-and-fuse-transport-endpoint-is-not-connected
#if auth is needed gcloud auth application-default login --no-browser and open link on local machine
#gcsfuse cedo_model_data storagebucket

#rebuild and load the package
devtools::load_all() # restarts and loads

# List of packages for session
.packages = c("import","dplyr") #need to be loaded

# Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])

# Load packages into session 
lapply(.packages, require, character.only=TRUE)

#use this for packages
#specifies what functions to import
# https://cran.r-project.org/web/packages/import/vignettes/import.html
import::from(dplyr, mutate, distinct, select, group_by, summarise, keep_when = filter, rename)

#to fix access with google
#https://github.com/orgs/community/discussions/39210

#to fix access issues
#https://www.namehero.com/blog/solve-temporary-failure-in-name-resolution-error-in-linux/
#sudo resolvectl flush-caches
#sudo systemd-resolve --flush-caches

system("sudo chmod -R a+rwx ~/", wait = TRUE)

devtools::install_github("markpayneatwork/ClimateOperators")

```

Get catch from official record. 
Filter out catch for artisanal fisheries and eliminate aquaculture production.
Uses table with corrected names to standarize names of locations and fishery offices and correct incorrect state names.
Combines catch files for all years and selects landing site, state, fishery office, species, weight, and catch value.
Resulting data is imported as catch rmd into the package

```{r get catch data, echo=TRUE}

locs.pesca <- clean_locs_pesca()

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

#Currently only using data 2018-2023, other data have too many inconsistencies
#catch.files <- list.files(path = here::here("data-raw","original_fisheries_data"), full.names = TRUE)
catch.files <- list.files(path = here::here("data-raw","original_fisheries_data"), full.names = TRUE, pattern = "2018|2019|2020|2021|2022|2023")

no.catchfiles <- 1:length(catch.files)

#NOTE: There are 16,531 records between 2000-2004 with 0 price for catch

#parallel::mclapply(no.catchfiles, get_catch, catch.files, coast.selecc, locs.pesca)
#use lapply for debugging
lapply(no.catchfiles, get_catch, catch.files, coast.selecc, locs.pesca)

catch.files.clean <- list.files(path = here::here("data-raw","fisheries_data"), full.names = TRUE)
no.catchfiles.clean <- 1:length(catch.files.clean)

catchdata <- lapply(no.catchfiles.clean, clean_catch, catch.files.clean) %>% 
  dplyr::bind_rows() %>% 
  mutate(catch_id = paste0("catch_",1:nrow(.)))

usethis::use_data(catchdata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(catchdata)
cat(data.description, file=here::here("R",paste0("data-","catchdata")))

#rebuild and load the package
devtools::load_all() # restarts and loads


```


Gets list of fishery offices from catch records
Uses standarized catch data to get unique fishery offices and georeference them
```{r Get fishery offices}

locs.pesca <- clean_locs_pesca()

#get unique georeferenced rnp codes

data("inegigc")
data("catchdata")

fishery.offices.list <- get_fishery_office_catch(catchdata) 

#fishery office list for 2024 is available here https://www.conapesca.gob.mx/work/sites/cona/dgppe/Subdelegaciones_de_Pesca.pdf
#the La Reforma office is in the Municipality of Angostura and has CVE_LOC = 250020093
#there are locations named La Reforma sites in Sinaloa so had to be extracted manually

fishery.office.table <- fishery.offices.list %>% 
  dplyr::bind_rows() %>% 
  dplyr::arrange(NOM_LOC,NOM_ENT)

fishery.spatial <- sf::st_as_sf(fishery.office.table, coords = c("deci_lon", "deci_lat"), crs = 4326) %>% 
  sf::st_transform(4269)

fishery.spatial %>% 
  sf::st_write(here::here("outputs","fishery_offices.shp"), append=FALSE)

readr::write_csv(fishery.office.table, here::here("outputs","fishery_offices.csv"))

fisheryoffice <- fishery.office.table
  
usethis::use_data(fisheryoffice, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(fisheryoffice)
cat(data.description, file=here::here("R",paste0("data-","fisheryoffice")))

#rebuild and load the package
devtools::load_all() # restarts and loads




```

Get locations from catch records
```{r Get locations from catch, echo=TRUE}

locs.pesca <- clean_locs_pesca()

data("catchdata")

data("inegigc")

catch.locations.data <- get_locations_catch(catchdata) 

readr::write_csv(catch.locations.data, here::here("outputs","catch_locations.csv"))

catchlocations <- catch.locations.data %>% 
  dplyr::bind_rows() 

usethis::use_data(catchlocations, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(catchlocations)
cat(data.description, file=here::here("R",paste0("data-","catchlocations")))

#rebuild and load the package
devtools::load_all() # restarts and loads


```


Create table georeferrenced locations from INEGI & Google and add distance to nearest fishery office

```{r location table, echo=TRUE}

locs.pesca <- clean_locs_pesca()

correct.ent <- locs.pesca %>%
    select(-NOM_LOC_REV)

correct.loc <- locs.pesca %>%
    select(-NOM_ENT_REV) 

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

#will extract all locations within states, not only coastal areas
#only run once will create attached data
#get_inegi(ents.coast.selecc)

data("inegigc")

locs.census <- inegigc %>% 
   select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)
  
#get georeferenced locations from inegi
locs.inegi <- googlesheets4::read_sheet("1Nww_0aSf2yQv9YqK8LqmJmIM2wwpHZCSpG1NcGfIfHQ", sheet = "busqueda_inegi") %>% 
  dplyr::rename(deci_lat=latitud, deci_lon=longitud) %>% 
  keep_when(utilizar == 1) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
  select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)

#test for duplicates 
dupes.inegi <- locs.inegi %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1) %>% 
  keep_when(dupe == TRUE)

#get georeferenced locations from google
locs.google <- googlesheets4::read_sheet("1Nww_0aSf2yQv9YqK8LqmJmIM2wwpHZCSpG1NcGfIfHQ", sheet = "busqueda_google") %>% 
  dplyr::rename(NOM_LOC=NOM_LOC_REV, deci_lat=lat, deci_lon=lon) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
  keep_when(utilizar == 1) %>% 
  mutate(deci_lon=as.numeric(deci_lon), deci_lat=as.numeric(deci_lat)) %>% 
  select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)

#test for duplicates 
dupes.google <- locs.google %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1)

#Atlas fishing locations Pacific
locs.pesca.pacifico <- googlesheets4::read_sheet("1KxO30FqYed655frJDiLobx3N-slFNpNLpHnbxE4cLcI") %>% 
  dplyr::rename(NOM_ENT= ESTADO, NOM_LOC = LOCALIDAD_REVISADA) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  select(-LOCALIDAD_ORIGINAL, -ID) %>% 
  dplyr::rename(deci_lat =LATITUD, deci_lon=LONGITUD) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
   select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)


#test for duplicates 
dupes.pacifico <- locs.pesca.pacifico %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1)

#combine all locations
georef.locs <- dplyr::bind_rows(locs.census, locs.inegi, locs.google, locs.pesca.pacifico) %>% 
  distinct(NOM_ENT, NOM_LOC, deci_lat, deci_lon) 

#make spatial

georef.locs.sp <- georef.locs %>% 
  sf::st_as_sf(coords = c("deci_lon", "deci_lat"), crs = 4326) %>% 
  sf::st_transform(4269)

#Add distance to nearest fishery office to help resolve duplicates when georeferincing a catch 

fishery.offices <- sf::st_read(here::here("outputs","fishery_offices.shp")) %>%   sf::st_transform(crs = 4326) %>% 
  sf::st_centroid() 

num.georef.locs <- 1:nrow(georef.locs.sp)

NumberOfCluster <- parallel::detectCores()  #- 1

georef.locs.dist <- parallel::mclapply(num.georef.locs, get_distance, fishery.offices, mc.cores = NumberOfCluster) %>% 
  dplyr::bind_rows() 


georeflocskm <- georef.locs.dist %>% 
  mutate(distance_km = distance / 1000)

readr::write_csv(georeflocskm, here::here("outputs","georef_locations.csv"))

usethis::use_data(georeflocskm, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(georeflocskm)
cat(data.description, file=here::here("R",paste0("data-","georeflocskm")))

#rebuild and load the package
devtools::load_all() # restarts and loads

```


Georeference unique locations
```{r Georeference locations, echo=TRUE}

data("catchlocations")

data("georeflocskm")


georefcatchlocs <- georef_catch(catchlocations, georeflocskm)
  
usethis::use_data(georefcatchlocs, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(georefcatchlocs)
cat(data.description, file=here::here("R",paste0("data-","georefcatchlocs")))

#rebuild and load the package
devtools::load_all() # restarts and loads

```

Get permit information from the Registro Nacional de Pesca y Acuacultura RNPA and georeference
https://www.gob.mx/conapesca/documentos/registro-nacional-de-pesca-y-acuacultura-rnpa

```{r Get permit data, echo=TRUE, message=TRUE, warning=TRUE}

data("georeflocskm")

locs.pesca <- clean_locs_pesca()

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

inegi.cost.comm.all  <- readr::read_csv(here::here("inputs","all_coastal_communities_INEGI.csv")) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_MUN = stringi::stri_trans_general(str = NOM_MUN, id = "Latin-ASCII"), NOM_MUN = toupper(NOM_MUN)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  keep_when(NOM_ENT %in% coast.selecc)

inegi.cost.mun <- inegi.cost.comm.all %>% 
  distinct(NOM_MUN, CVE_MUN)

permit.files <- list.files(path = here::here("data-raw","permit_data"), full.names = TRUE)
no.permitfiles <- 1:length(permit.files)

permit.data <- lapply(no.permitfiles, extract_permits, permit.files, coast.selecc, locs.pesca, georeflocskm)

permitdata <- permit.data %>% 
  dplyr::bind_rows() %>% 
  distinct(rnp_code, NOM_LOC, NOM_ENT, deci_lat, deci_lon, fishery_office, fishery_office_ent, distance_km)


usethis::use_data(permitdata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(permitdata)
cat(data.description, file=here::here("R",paste0("data-","permitdata")))

#rebuild and load the package
devtools::load_all() # restarts and loads

readr::write_csv(permitdata, here::here("outputs","rnp_fishery_permits.csv"))


```


Georeference all catch data
```{r Georeference catch, echo=TRUE}

data("georefcatchlocs")
data("catchdata")
data("permitdata")

if(file.exists("outputs/gpkg_grid.gpkg")) file.remove("outputs/gpkg_grid.gpkg")

catch.plot.inter <- georef_catch_data(georefcatchlocs, catchdata)

#plotly::ggplotly(catch.plot.inter)

catch.plot.inter

```

Create grid for environmental data
```{r Create grid, echo=TRUE}

gc.grid.plot <- make_grid()

gc.grid.plot 
```

Download GLORYS data
```{r download GLORYS, eval=FALSE, include=FALSE}
# Script to download GLORYS data directly from Copernicus Servers
# Requires a Copernicus marine service account
# https://data.marine.copernicus.eu/products

#mount storage bucket

#if cannot connect to servers
# sudo systemd-resolve --flush-caches
# sudo resolvectl flush-caches
# sudo nmcli general reload dns-full
#to mount bucket 
#https://cloud.google.com/storage/docs/gcsfuse-mount#authenticate_requests
#https://medium.com/google-cloud/how-to-use-gcs-bucket-with-gcp-vm-using-gsutil-and-mounting-as-a-filesystem-9139b812db84
#this can also be run if you lose write access to the storage bucket
#gcloud auth application-default login --no-launch-browser
#gcsfuse cedo_model_data storagebucket


#system("sudo chmod -R a+rwx ~/storagebucket/GLORYS", wait = TRUE)

#Install and load dependencies
#system("pip install motuclient==1.8.4")
#install.packages("reticulate")
library(reticulate)
library(lubridate)
#only need to create environment first time
#install_python() 
#virtualenv_create(envname = "copernicus")
#virtualenv_install("copernicus", packages = c("copernicusmarine"))
use_virtualenv("copernicus", required = TRUE)
#py_install("copernicusmarine")

cm <- import("copernicusmarine")

#Specify lat/lon/depth bounds
#Gulf of California 
#-116.1407,20.44,-105.1029,32.0482
#Southeast limit at Cabo Corrientes 20.43386153865933, -105.69209230968688

min_lon <- -116.1407
max_lon <- -105.1000
min_lat  <- 20.4339
max_lat  <- 32.0482

#only needed once unless credentials need to be updated
#cm$login("hmorzarialuna", "#esBmZe7P_H7r_i")

#Set directories
out_dir = paste0("~/storagebucket/GLORYS/monthly")

#Specify GLORYS file from manual https://catalogue.marine.copernicus.eu/documents/PUM/CMEMS-GLO-PUM-001-030.pdf

#GLORYS monthly available time series is 04/12/1991 to 31/06/2021
data.set.id <- "cmems_mod_glo_phy_my_0.083deg_P1M-m"
var.id <- "thetao"
#set time period
#start.time.interval <- seq(as.Date("2018-01-01"), length=42, by="1 month") 
#end.time.interval <- seq(as.Date("2018-02-01"), length=42, by="1 month") - 1
start.time.interval <- seq(as.Date("2018-01-01"), length=8, by="1 month") 
end.time.interval <- seq(as.Date("2018-02-01"), length=8, by="1 month") - 1
#set length of time period
entry.interval <- 1:length(start.time.interval)
#download data
lapply(entry.interval, get_monthlyglorys, start.time.interval, end.time.interval, min_lon, max_lon, min_lat, max_lat, out_dir, data.set.id)


#GLORYS monthly available time series is 01/07/2021 to M-4 (current month - 4), need to get data 01/07/2021 12/31/2023
data.set.id <- "cmems_mod_glo_phy_myint_0.083deg_P1M-m"
var.id <- "thetao"
#set time period
start.time.interval <- seq(as.Date("2021-07-01"), length=30, by="1 month") 
end.time.interval <- seq(as.Date("2021-08-01"), length=30, by="1 month") - 1
#set lenght of time period
entry.interval <- 1:length(start.time.interval)
#download data
lapply(entry.interval, get_monthlyglorys, start.time.interval, end.time.interval, min_lon, max_lon, min_lat, max_lat, out_dir, data.set.id)

#Get GLORYS data prior to 2018 for climatology
#GLORYS monthly available time series is 04/12/1991 to 31/06/2021
data.set.id <- "cmems_mod_glo_phy_my_0.083deg_P1M-m"
var.id <- "thetao"
#set time period
start.time.interval <- seq(as.Date("1994-01-01"), length=288, by="1 month") 
end.time.interval <- seq(as.Date("1994-02-01"), length=288, by="1 month") - 1
#set length of time period
entry.interval <- 1:length(start.time.interval)
#download data
lapply(entry.interval, get_monthlyglorys, start.time.interval, end.time.interval, min_lon, max_lon, min_lat, max_lat, out_dir, data.set.id)

#1993-2016 climatology
data.set.id <- "cmems_mod_glo_phy_my_0.083deg-climatology_P1M-m"
out_dir = paste0("~/storagebucket/GLORYS/climatology")

#download data
get_climatologyglorys(min_lon, max_lon, min_lat, max_lat, out_dir, data.set.id)

#To get forecast data from Global Ocean Physics Analysis and Forecast
#https://data.marine.copernicus.eu/product/GLOBAL_ANALYSISFORECAST_PHY_001_024/description

#To get biogeochemistry data
#https://catalogue.marine.copernicus.eu/documents/PUM/CMEMS-GLO-PUM-001-029.pdf
#data.set.id <- "cmems_mod_glo_bgc_my_0.083deg_P1M-m"
#var.id <- "chl"

```


map GLORYS data
```{r Map GLORYS data, echo=TRUE}

catch.pacific.spatial <- sf::st_read(dsn="outputs/gpkg_grid.gpkg", layer='goc_catch')

#select example GLORYS data
glorys.file <- paste0("/home/atlantis/storagebucket/GLORYS/monthly/","cmems_mod_glo_phy_myint_0.083deg_P1M-m_thetao_116.08W-105.17W_20.50N-32.00N_0.49-453.94m_2023-08-01.nc")
this.terra.rast<- get_nc_description(glorys.file)
terra::plot(this.terra.rast)

#esm.file  <- here("delta_method_final_outputs","GFDL-CM4_delta_thetao_2050.nc")
#glorys.esm.file  <- here("GLORYS_ESM_final","GFDL-CM4_glorys_thetao_2050.nc")

glorys.nc  <-  tidync::tidync(glorys.file) %>%
  tidync::hyper_tibble() %>%
  dplyr::filter(depth == min(depth))

# esm.nc = tidync(esm.file)%>%
#   hyper_tibble()%>%
#   filter(lev == min(lev))
# 
# glorys.esm.nc = tidync(glorys.esm.file)%>%
#   hyper_tibble()%>%
#   filter(lev == min(lev))

catch.geo.pacific <- catch.pacific.spatial %>% 
    dplyr::mutate(longitude = sf::st_coordinates(.)[,1],
                  latitude = sf::st_coordinates(.)[,2]) %>% 
    dplyr::as_tibble() 

glorys.goc <- ggplot2::ggplot(data = glorys.nc, ggplot2::aes(x= longitude, y = latitude, color = thetao))+
  ggplot2::geom_point()+
#  ggplot2::annotation_map(mapdata::map('worldHires'),fill = 'grey70',alpha=0.3) +
  ggplot2::labs(title="2023-08-01",
                color = "Sea temperature") +
  ggplot2::geom_point(data = catch.geo.pacific, color = "black") 
  

glorys.goc
```

Aggregate depth bins to match projections of CMIP6 data

```{r}
glorys.files <- list.files(path = "/home/atlantis/storagebucket/GLORYS/monthly", pattern="*01.nc$", full.names = TRUE)

#need to aggregate initial depth bins because projected future data starts at 6 m and it is not possible to interpolate to smaller bins
ClimateOperators::cdo("zaxisdes",glorys.files[1])  #shows depth, z axis
ClimateOperators::cdo("sinfon",glorys.files[1])  #shows depth, z axis
    agg.lev <- 7.92956 #will select all depths below this one for aggregation
    max.lev <- 34.43415 #max level for selecting

lapply(glorys.files,agg_depth_bin, agg.lev, max.lev)

```


Get temperature data for each point

```{r Get temp data per point, echo=FALSE}
#Get temperature for each point from closest raster cell 
#uses disaggregated temperature files

this.depth <- 10 #Most species are not caught deeper
# gc.ocean.grid <- sf::st_read(dsn="outputs/goc_grid.shp") %>% 
#   dplyr::mutate(id_pol=1:nrow(.))

catch.pacific.spatial <- sf::st_read(dsn="outputs/gpkg_grid.gpkg", layer='goc_catch') %>% 
  dplyr::mutate(catch_id = 1:nrow(.)) %>% 
    mutate(month_no = dplyr::case_when(
      month  == "DICIEMBRE" ~ "12",
      month == "NOVIEMBRE" ~ "11",
      month == "OCTUBRE" ~ "10",
      month == "SEPTIEMBRE" ~ "09",
      month == "AGOSTO" ~ "08",
      month == "JULIO" ~ "07",
      month == "JUNIO" ~ "06",
      month == "MAYO" ~ "05",
      month == "ABRIL" ~ "04",
      month == "MARZO" ~ "03",
      month == "FEBRERO" ~ "02",
      month == "ENERO" ~ "01"))
#Southeast limit at Cabo Corrientes 20.43386153865933, -105.69209230968688

catch.geo <- catch.pacific.spatial %>% 
    dplyr::mutate(deci_lon = sf::st_coordinates(.)[,1],
                  deci_lat = sf::st_coordinates(.)[,2]) %>% 
    dplyr::as_tibble() %>% 
    select(-geom) 

years.data <- rep(seq(from=2018, to=2023), each = 12)
start.time.interval <- paste(years.data,c("01","02","03","04","05","06","07","08","09","10","11","12"), sep = "-")

time.interval <- 1:length(start.time.interval)

glorys.files <- list.files(path = "/home/atlantis/storagebucket/GLORYS/monthly_original", pattern= "*.nc",full.names = TRUE)

NumberOfCluster <- parallel::detectCores()  #- 1

datapoints <- parallel::mclapply(time.interval, get_nearestpointtemp, this.depth, catch.pacific.spatial, glorys.files, catch.geo, start.time.interval, out.dir = "/home/atlantis/storagebucket/GLORYS/pointtemp", mc.cores = NumberOfCluster)


plot.catch <- ggplot2::ggplot(data = catch.pacific.spatial, ggplot2::aes(x= year, y = landed_w_kg, color = species))+
  ggplot2::geom_line()+
  ggplot2::labs(title="Catch time series") +
  ggplot2::facet_wrap(.~species)

```

Get temperature for convex hull created by cluster points
```{r get cluster temp}

catch.pacific.spatial <- sf::st_read(dsn="outputs/gpkg_grid.gpkg", layer='goc_catch') %>% 
   dplyr::mutate(deci_lon = sf::st_coordinates(.)[,1],
                  deci_lat = sf::st_coordinates(.)[,2])

cluster.locs <- data.table::fread("~/storagebucket/IA-Input-data/data_clas.csv") %>% 
  dplyr::filter(!is.na(Cluster_Label)) %>% 
  dplyr::select(year:deci_lat, Cluster_Label) %>% 
  sf::st_as_sf(coords = c("deci_lon", "deci_lat"), crs = 4269) %>% 
  dplyr::mutate(deci_lon = sf::st_coordinates(.)[,1],
                deci_lat = sf::st_coordinates(.)[,2])

#use to visualize clusters
gc.cluster.plot <- cluster.locs %>% 
  mutate(clusters=as.factor(Cluster_Label)) %>% 
    ggplot2::ggplot() +
    ggplot2::geom_sf() +
    ggplot2::geom_sf(ggplot2::aes(color = clusters), alpha=0.4) +
   ggplot2::labs(x = "Longitude",
                  y="Latitude",
                  title = "Georeferenced catch data",
                  subtitle = "Grid 0.083° × 0.083° degree")
          
gc.cluster.plot    
num.clusters <- unique(cluster.locs$Cluster_Label)

years.data <- rep(seq(from=2018, to=2023), each = 12)
start.time.interval <- paste(years.data,c("01","02","03","04","05","06","07","08","09","10","11","12"), sep = "-")

time.interval <- 1:length(start.time.interval)

glorys.files <- list.files(path = "/home/atlantis/storagebucket/GLORYS/monthly", pattern="._setlev.nc",full.names = TRUE)

NumberOfCluster <- parallel::detectCores()  #- 1

 parallel::mclapply(time.interval, get_meanclustertemp, glorys.files, cluster.locs, num.clusters, time.interval, start.time.interval, out.dir = "/home/atlantis/storagebucket/GLORYS/clustertemp", mc.cores = NumberOfCluster)


```

Get CMIP6 files from repository and long-term storage
```{r get CMIP6 ESM files}
#Get ESM model files for climatology from cedo archive
#store in storage bucket /ESM_CMIP6

#sh files stored in localidadesarribo/inputs
system("sudo chmod u+x wget_script_2024-6-27_17-53-18.sh")
system("sudo bash ./wget_script_2024-6-27_17-53-18.sh -s") # -s option allows certificate free download
#see https://esgf.github.io/esgf-user-support/user_guide.html#data-search-and-download
#Downloads files stored in Azure blob storage
system("sudo chmod + esm_files.sh")
system("sudo sh ./esm_files.sh")

system("sudo chmod + esm_future.sh")
system("sudo sh ./esm_future.sh")

#to run CDO, there is an error in Ubuntu repositories, see https://code.mpimet.mpg.de/boards/1/topics/12307?r=13297#message-13297
#removed and reinstalled CDO using Conda https://github.com/conda-forge/cdo-feedstock

```


Regrid files to assign lat lon as dimensions instead of variables
```{r regrid files, include=FALSE}

#CDO manual https://users.rcc.uchicago.edu/~davidkelly999/cdo.pdf
# https://github.com/markpayneatwork/ClimateOperators
esm.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/historical", pattern="*2.nc", full.names = TRUE)
this.file <- esm.files[2]
this.terra.rast<- get_nc_description(this.file)
terra::plot(this.terra.rast)

#NumberOfCluster <- parallel::detectCores()  - 1

lapply(esm.files,correct_grid, this.var="tos")

#parallel::mclapply(esm.files,correct_grid, this.var="tos", mc.cores = NumberOfCluster)

esm.files.future <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/future", pattern="*2.nc", full.names = TRUE)

#parallel::mclapply(esm.files.future, correct_grid, this.var="tos", mc.cores = NumberOfCluster)

lapply(esm.files.future, correct_grid, this.var="tos")

```

Hyperslab nc files to extract study area
```{r spatial crop, include=FALSE}

esm.coord.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/historical", pattern="*_coord.nc", recursive=TRUE, full.names = TRUE)
this.terra.rast<- get_nc_description(esm.coord.files[2])
terra::plot(this.terra.rast)


#-118.0008537881,19.9871469981,-104.9710686319,32.1040488772
lapply(esm.coord.files, hyperslab_spatial_area, min.lon=-118.0008537881, max.lon=-104.9710686319, min.lat= 19.9871469981, max.lat= 32.1040488772)

esm.coord.future.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/future", pattern="*_coord.nc", full.names = TRUE)

lapply(esm.coord.future.files, hyperslab_spatial_area, min.lon=-118.0008537881, max.lon=-104.9710686319, min.lat= 19.9871469981, max.lat= 32.1040488772)


```

Merge together files from multiple decades
```{r merge files, include=FALSE}

esm.slab.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/historical", pattern="*_slab.nc", full.names = TRUE)
esm.slab.future.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/future", pattern="*_slab.nc", full.names = TRUE)

this.terra.rast<- get_nc_description(esm.slab.files[2])
terra::plot(this.terra.rast)

this.terra.rast.future<- get_nc_description(esm.slab.future.files[2])
terra::plot(this.terra.rast.future)

#historical files will be 1990-2024, uses two files from future projections
#realizations to loop through
#can only use realization r1 and r2 because they are the only future realizations
rel.list <- paste0("r",1:2,"i1p1f1")
merged.dir <- "/home/atlantis/storagebucket/ESM_CMIP6/thetao/merged_files/"
dir.create(merged.dir)
#future data are the years from future projections that will be added to historical to match catch data or removed from future data
#data period is entire period covered

lapply(rel.list, merge_files, esm.slab.files, esm.slab.future.files, merged.dir, data.type = "historical", data.period=seq(1990,2024, by=5), data.range = c("199001","202412"))

#50 year projection dataset
lapply(rel.list, merge_files, esm.slab.files, esm.slab.future.files, merged.dir, data.type = "future", data.period=seq(2025,2070, by=5), data.range = c("202501","207412"))

```

Hyperslab of depth layers
```{r}
#select 0-30 to match selcted depths in GLORYS files, data goes to 5000m but we don't need that
esm.merge.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/merged_files", pattern="*_merged.nc", full.names = TRUE)
this.terra.rast<- get_nc_description(esm.merge.files[2])
terra::plot(this.terra.rast)


#nco uses indexing of depth layers
ini.depth = 1
end.depth = 4

#lev = 6, 17, 27, 37, 47, 57, 68.5, 82.5, 100, 122.5, 150, 182.5, 220, 262.5, 
#    310, 362.5, 420, 485, 560, 645, 740, 845, 960, 1085, 1220, 1365, 1525, 
#    1700, 1885, 2080, 2290, 2525, 2785, 3070, 3395, 3770, 4195, 4670, 5170, 
#    5720 ;

lapply(esm.merge.files,hyperslab_depth,ini.depth, end.depth)

esm.depth.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/merged_files", pattern="*_depth.nc", full.names = TRUE)

system(paste0("cdo sinfon ",esm.depth.files[3]))
```

Calculate ESM climatology
```{r calculate ESM climatology, include=FALSE}
#Calculate 30 or 50 year climatology

depth.hist.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/merged_files", pattern="1990.+depth", full.names = TRUE)
depth.future.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/merged_files", pattern="2025.+depth", full.names = TRUE)
this.terra.rast<- get_nc_description(depth.hist.files[2])
terra::plot(this.terra.rast)

this.terra.rast.future <- get_nc_description(depth.future.files[2])
terra::plot(this.terra.rast.future)


rel.list <- paste0("r",1:2,"i1p1f1")
clim.dir <- "/home/atlantis/storagebucket/ESM_CMIP6/thetao/clim_files/"
dir.create(clim.dir)
delta.dir <- "/home/atlantis/storagebucket/ESM_CMIP6/thetao/delta_files/"
dir.create(delta.dir)

delta.future.file<- lapply(rel.list, cal_clim, depth.hist.files, depth.future.files)
 
delta.rast.1 <- terra::rast(delta.future.file[[1]])
terra::plot(delta.rast.1)

delta.rast.2 <- terra::rast(delta.future.file[[2]])
terra::plot(delta.rast.2)

  
```

Interpolate anomalies to full extent of GLORYS grid

```{r}

esm.delta.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/delta_files", pattern="*delta.nc", full.names = TRUE)
this.delta.file <- esm.delta.files[2]
this.terra.rast<- get_nc_description(this.delta.file)
terra::plot(this.terra.rast)
delta.path <- "~/storagebucket/ESM_CMIP6/thetao/delta_files/"

#pick one GLORYS file for grid extraction
glorys.nc <- "~/storagebucket/GLORYS/monthly/cmems_mod_glo_phy_my_0.083deg_P1M-m_thetao_116.08W-105.17W_20.50N-32.00N_0.49-453.94m_1997-02-01_setlev.nc"
system(paste0("cdo infon ",glorys.nc),wait = TRUE)
terra::plot(this.terra.rast)

lapply(esm.delta.files, interpolate_delta, glorys.nc, delta.path)

esm.interp.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/delta_files", pattern="*interp.nc", full.names = TRUE)

this.terra.rast.1<- get_nc_description(esm.interp.files[1])
terra::plot(this.terra.rast.1)

this.terra.rast.2<- get_nc_description(esm.interp.files[2])
terra::plot(this.terra.rast.2)
  
```
Climatology GLORYS
```{r merge clim glorys, include=FALSE}

#list monthly files to merge and calculate climatology
glorys.monthly.files <- list.files(path = "~/storagebucket/GLORYS/monthly", pattern="*setlev.nc", full.names = TRUE)

#climatology will be 1994-2023

merged.dir <- "/home/atlantis/storagebucket/GLORYS/merged_files/"
dir.create(merged.dir)
#future data are the years from future projections that will be added to historical to match catch data or removed from future data
glorys.name <- "cmems_mod_glo_phy_my_0.083deg_P1M-m_thetao_116.08W-105.17W_20.50N-32.00N_0.49-453.94m_193001-202412_merged.nc"

clim.glorys.file <- merge_glorys(glorys.monthly.files, merged.dir, glorys.name)

this.terra.rast<- get_nc_description(clim.glorys.file)
terra::plot(this.terra.rast)

```



Add the grid-interpolated ESM deltas to the GLORYS climatology
```{r}

glorys.clim.files <- list.files(path = "/home/atlantis/storagebucket/GLORYS/merged_files", pattern="*clim.nc", full.names = TRUE)
esm.inter.files <- list.files(path = "/home/atlantis/storagebucket/ESM_CMIP6/thetao/delta_files", pattern = "*minterp.nc", full.names = TRUE)

this.terra.rast<- get_nc_description(glorys.clim.files[1])
terra::plot(this.terra.rast)

this.terra.rast<- get_nc_description(esm.inter.files[1])
terra::plot(this.terra.rast)

rel.list <- paste0("r",1:2,"i1p1f1")

delta.dir <- "/home/atlantis/storagebucket/ESM_CMIP6/thetao/delta_files"

lapply(rel.list, interp_clim, esm.inter.files, glorys.clim.files, delta.dir, type ="add", timestep="month")


 
```
map monthly future oceanographic projections on the GLORYS grid
```{r Map future data, echo=TRUE}

catch.pacific.spatial <- sf::st_read(dsn="outputs/gpkg_grid.gpkg", layer='goc_catch')

#select example future data
future.files <- list.files(path="/home/atlantis/storagebucket/ESM_CMIP6/thetao/delta_files",pattern=".projfut.", full.names = TRUE)
this.future.file <- future.files[55]
this.terra.rast<- get_nc_description(this.future.file)
terra::plot(this.terra.rast)

#esm.file  <- here("delta_method_final_outputs","GFDL-CM4_delta_thetao_2050.nc")
#glorys.esm.file  <- here("GLORYS_ESM_final","GFDL-CM4_glorys_thetao_2050.nc")

future.nc  <-  tidync::tidync(this.future.file) %>%
  tidync::hyper_tibble() %>%
  dplyr::filter(lev == min(lev))

# esm.nc = tidync(esm.file)%>%
#   hyper_tibble()%>%
#   filter(lev == min(lev))
# 
# glorys.esm.nc = tidync(glorys.esm.file)%>%
#   hyper_tibble()%>%
#   filter(lev == min(lev))

catch.geo.pacific <- catch.pacific.spatial %>% 
    dplyr::mutate(longitude = sf::st_coordinates(.)[,1],
                  latitude = sf::st_coordinates(.)[,2]) %>% 
    dplyr::as_tibble() 

future.goc <- ggplot2::ggplot(data = future.nc, ggplot2::aes(x= longitude, y = latitude, color = thetao))+
  ggplot2::geom_point()+
#  ggplot2::annotation_map(mapdata::map('worldHires'),fill = 'grey70',alpha=0.3) +
  ggplot2::labs(title="Future temperature 2050",
                color = "Sea temperature") +
  ggplot2::geom_point(data = catch.geo.pacific, color = "black") 
  

future.goc
```

Get future temperature for convex hull created by cluster points
```{r get cluster temp fut}

catch.pacific.spatial <- sf::st_read(dsn="outputs/gpkg_grid.gpkg", layer='goc_catch') %>% 
   dplyr::mutate(deci_lon = sf::st_coordinates(.)[,1],
                  deci_lat = sf::st_coordinates(.)[,2])

cluster.locs <- data.table::fread("~/storagebucket/IA-Input-data/data_clas.csv") %>% 
  dplyr::filter(!is.na(Cluster_Label)) %>% 
  dplyr::select(year:deci_lat, Cluster_Label) %>% 
  sf::st_as_sf(coords = c("deci_lon", "deci_lat"), crs = 4269) %>% 
  dplyr::mutate(deci_lon = sf::st_coordinates(.)[,1],
                deci_lat = sf::st_coordinates(.)[,2])

#use to visualize clusters
gc.cluster.plot <- cluster.locs %>% 
  mutate(clusters=as.factor(Cluster_Label)) %>% 
    ggplot2::ggplot() +
    ggplot2::geom_sf() +
    ggplot2::geom_sf(ggplot2::aes(color = clusters), alpha=0.4) +
   ggplot2::labs(x = "Longitude",
                  y="Latitude",
                  title = "Georeferenced catch data",
                  subtitle = "Grid 0.083° × 0.083° degree")
          
gc.cluster.plot    
num.clusters <- unique(cluster.locs$Cluster_Label)

start.time.interval <- rep(seq(from=2025, to=2074), each = 1)

time.interval <- 1:length(start.time.interval)

future.files <- list.files(path="/home/atlantis/storagebucket/ESM_CMIP6/thetao/delta_files",pattern=".projfut.", full.names = TRUE)

NumberOfCluster <- parallel::detectCores()  #- 1

out.dir  <-  "/home/atlantis/storagebucket/ESM_CMIP6/futureclustertemp"
dir.create(out.dir)

parallel::mclapply(time.interval, get_meanclustertempfuture, future.files, cluster.locs, num.clusters, time.interval, start.time.interval, out.dir = "/home/atlantis/storagebucket/ESM_CMIP6/futureclustertemp", mc.cores = NumberOfCluster)


```

Interpolate with nco
```{r}
#https://acme-climate.atlassian.net/wiki/spaces/DOC/pages/754286611/Regridding+E3SM+Data+with+ncremap
#Supply ncremap with a data-template file (dat.nc) and a grid-file name (grd.nc). Since there are no input files to regrid, ncremap exits after inferring the grid-file
#ncremap -d dat.nc -g grd.nc


esm.delta.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/delta_files", pattern="*delta.nc", full.names = TRUE)
this.delta.file <- esm.delta.files[1]

glorys.nc <- "~/storagebucket/GLORYS/monthly/cmems_mod_glo_phy_my_0.083deg_P1M-m_thetao_116.08W-105.17W_20.50N-32.00N_0.49-453.94m_1997-02-01_setlev.nc"

map.file <- "~/storagebucket/ESM_CMIP6/thetao/delta_files/map_file.nc"
glorys.grid <- "~/storagebucket/ESM_CMIP6/thetao/delta_files/glorys_grid.nc" 
esm.grid <- "~/storagebucket/ESM_CMIP6/thetao/delta_files/esm_grid.nc" 


#inferr grid file from source ESM data file
system(paste0("ncremap -d ",this.delta.file," -g ",esm.grid), wait=TRUE)

#inferr grid file from target glorys data file
system(paste0("ncremap -d ",glorys.nc," -g ",glorys.grid), wait=TRUE)

#generate map file from grid files
#ncremap -s grd_src.nc -g grd_dst.nc -m map.nc
system(paste0("ncremap -s ",esm.grid," -g ",glorys.grid," -m ", map.file), wait=TRUE)

#name output interpolated file
this.inter.file <- gsub("delta.nc","interp.nc",this.delta.file)

#regrid using map file
#ncremap -m map.nc dat_src.nc dat_rgr.nc
system(paste0("ncks --rnr=0.0 --map=map_file.nc ",this.delta.file, " ",this.inter.file))

clim.rast <- terra::rast(this.inter.file)
terra::plot(clim.rast)
```


```{r}
##### IDW interpolation #####

#https://rdrr.io/github/rspatial/terra/man/interpolate.html
#https://nceas.github.io/oss-lessons/spatial-data-gis-law/4-tues-spatial-analysis-in-r.html

esm.interp.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/delta_files", pattern="*interp.nc", full.names = TRUE)
this.interp.file <- esm.interp.files[1]
terra.rast <- terra::rast(this.interp.file)

terra.points <- terra::as.points(terra.rast)

rast.data.frame <- terra::as.data.frame(terra.points)

gstat.rast <- gstat::gstat(id = "thetao", formula = thetao~1, locations = ~x+y, data=terra.rast, 
            nmax=terra.rast, set=list(idp = .5))

z <- terra::interpolate(terra.rast, gstat.rast, debug.level=0, index=1)


# establish an extent within which you want to interpolate
# -99/24 to -80/32.
x_range <- as.numeric(c(-117.8304, -104.81))  # min/max longitude of the interpolation area
y_range <- as.numeric(c(20.0495, 32.07921))  # min/max latitude of the interpolation area

# create an empty grid of values ranging from the xmin-xmax, ymin-ymax
grd <- expand.grid(x = seq(from = x_range[1],
                   to = x_range[2], 
                   by = 0.08333337),
                   y = seq(from = y_range[1], to = y_range[2], 
                   by = 0.08333337))  # expand points to grid
class(grd)

# points object
sp::coordinates(grd) <- ~x + y
# turn into a spatial pixels object
sp::gridded(grd) <- TRUE

sp::plot(grd)

mg <- gstat(id = "zinc", formula = zinc~1, locations = ~x+y, data=meuse, 
            nmax=7, set=list(idp = .5))
z <- interpolate(r, mg, debug.level=0, index=1)

```


Interpolation using Inverse Path Distance Weighing
IDW interpolation
```{r}

#convert netcdf file to csv 
esm.delta.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/delta_files", pattern="*delta.nc", full.names = TRUE)
this.delta.file <- esm.delta.files[2]

#GLORYS file for grid extraction
glorys.nc <- "~/storagebucket/GLORYS/monthly/cmems_mod_glo_phy_my_0.083deg_P1M-m_thetao_116.08W-105.17W_20.50N-32.00N_0.49-453.94m_1997-02-01_setlev.nc"

#regrid depths
this.intlevel.file <- gsub("_delta","_intlevel", this.delta.file)

  zlevs.glorys <- tidync::tidync(glorys.nc) %>% 
    tidync::activate("D3") %>% 
    tidync::hyper_tibble() %>% 
    dplyr::pull('sfc') %>% 
    paste(collapse=",")
  
system(paste0("cdo -intlevel,",zlevs.glorys," ",this.delta.file," ",this.intlevel.file), wait=TRUE)

get_nc_description(this.intlevel.file)

#hyperslab each depth
 zlevs.intlevel <- tidync::tidync(this.intlevel.file) %>% 
    tidync::activate("D4") %>% 
    tidync::hyper_tibble() %>% 
    dplyr::pull('lev') 
 
 for(eachlevel in 1:length(zlevs.intlevel)){
   
   this.zlevel <- zlevs.intlevel[eachlevel]
   
   this.depth.intlevel <- gsub("_intlevel",paste0("_intlevel_",eachlevel), this.intlevel.file)
   
   system(paste0("ncks -O -F -d"," lev,",this.zlevel," ", this.intlevel.file," ", this.depth.intlevel), wait=TRUE)
 this.raster.layer <- get_nc_description(this.depth.intlevel)
terra::plot(this.raster.layer)
   
#convert raster to xy data 

system(paste0("cdo -outputtab,date,lon,lat,value"," ", this.depth.intlevel," > outfile.txt"), wait = TRUE)
test <- readLines("outfile.txt")

ncin <- ncdf4::nc_open(this.depth.intlevel)
lon <- ncdf4::ncvar_get(ncin, "lon")
lat <- ncdf4::ncvar_get(ncin, "lat")
time <- ncdf4::ncvar_get(ncin, "time")
thetao <- ncdf4::ncvar_get(ncin, "thetao")
thetao.vec <- as.vector(thetao)
ncin.mat <- as.matrix(expand.grid(lon, lat, time))
thetao.mat <- data.frame(cbind(ncin.mat, thetao.vec))
thetao.mat$Var3 <- as.POSIXct(thetao.mat$Var3*175*365*2, origin='1850-1-1 00:00:00')
 }
    

this.terra.rast<- get_nc_description(this.delta.file)


system(paste0("cdo -outputtab,time,lat:6,lon:6,lev ",this.delta.file," >> outfile.csv"), wait= TRUE)

raster.data <- readr::read_delim("outfile.csv",delim=" ")
```


Tried running this but it almost fills RAM memory
```{r IDW, eval=FALSE, include=FALSE}
#Interpolation similar to splines
#https://cran.r-project.org/web/packages/ipdw/vignettes/ipdw2.html
#lat.lon.raster <- data.frame(xyFromCell(raster.nc.rotate.ext, 1:ncell(raster.nc.rotate.ext)))
#vals.raster <- getValues(raster.nc.rotate.ext)

esm.delta.files <- list.files(path = "~/storagebucket/ESM_CMIP6/thetao/delta_files", pattern="*delta.nc", full.names = TRUE)

this.delta.file <- esm.delta.files[1]

this.rast.file <- terra::rast(this.delta.file)

layer.subset <- terra::subset(this.rast.file,1)
point.layer <- terra::as.points(layer.subset) %>% 
  sf::st_as_sf(.) %>% 
  sf::st_transform(4485)

coast.pol <- sf::st_read(here::here("data-raw","SHP","Golfo_california_wetland_poly_WGS84.shp")) %>% 
  sf::st_transform(4485)

costras <- ipdw::costrasterGen(point.layer, coast.pol, extent = "pnts",
  projstr = raster::projection(coast.pol))

#costras <- costrasterGen(pnts, pols, extent = "pnts", projstr = raster::projection(pols))

```



