---
title: "Goreference landing sites for fishery logs"
output: html_notebook
---


```{r load package, include=FALSE}
#rebuild and load the package
devtools::load_all() # restarts and loads

# List of packages for session
.packages = c("import")

# Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])

# Load packages into session 
lapply(.packages, require, character.only=TRUE)

#use this for packages
#specifies what functions to import
# https://cran.r-project.org/web/packages/import/vignettes/import.html
import::from(dplyr, mutate, distinct, select, group_by, summarise, keep_when = filter, rename)

#to fix access with google
#https://github.com/orgs/community/discussions/39210

```

Get catch from official record. 
Filter out catch for artisanal fisheries and eliminate aquaculture production.
Uses table with corrected names to standarize names of locations and fishery offices and correct incorrect state names.
Combines catch files for all years and selects landing site, state, fishery office, species, weight, and catch value.
Resulting data is imported as catch rmd into the package

```{r get catch data, include=FALSE}

locs.pesca <- clean_locs_pesca()

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

#Currently only using data 2018-2023, other data have too many inconsistencies
#catch.files <- list.files(path = here::here("data-raw","original_fisheries_data"), full.names = TRUE)
catch.files <- list.files(path = here::here("data-raw","original_fisheries_data"), full.names = TRUE, pattern = "2018|2019|2020|2021|2022|2023")

no.catchfiles <- 1:length(catch.files)

#NOTE: There are 16,531 records between 2000-2004 with 0 price for catch

#parallel::mclapply(no.catchfiles, get_catch, catch.files, coast.selecc, locs.pesca)
#use lapply for debugging
lapply(no.catchfiles, get_catch, catch.files, coast.selecc, locs.pesca)

catch.files.clean <- list.files(path = here::here("data-raw","fisheries_data"), full.names = TRUE)
no.catchfiles.clean <- 1:length(catch.files.clean)

catchdata <- lapply(no.catchfiles.clean, clean_catch, catch.files.clean) %>% 
  dplyr::bind_rows() %>% 
  mutate(catch_id = paste0("catch_",1:nrow(.)))

usethis::use_data(catchdata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(catchdata)
cat(data.description, file=here::here("R",paste0("data-","catchdata")))

#rebuild and load the package
devtools::load_all() # restarts and loads


```


Gets list of fishery offices from catch records
Uses standarized catch data to get unique fishery offices and georeference them
```{r}

locs.pesca <- clean_locs_pesca()

#get unique georeferenced rnp codes

data("inegigc")
data("catchdata")

fishery.offices.list <- get_fishery_office_catch(catchdata) 

#fishery office list for 2024 is available here https://www.conapesca.gob.mx/work/sites/cona/dgppe/Subdelegaciones_de_Pesca.pdf
#the La Reforma office is in the Municipality of Angostura and has CVE_LOC = 250020093
#there are locations named La Reforma sites in Sinaloa so had to be extracted manually

fishery.office.table <- fishery.offices.list %>% 
  dplyr::bind_rows() %>% 
  dplyr::arrange(NOM_LOC,NOM_ENT)

fishery.spatial <- sf::st_as_sf(fishery.office.table, coords = c("deci_lon", "deci_lat"), crs = 4326) 

fishery.spatial %>% 
  sf::st_write(here::here("outputs","fishery_offices.shp"), append=FALSE)

readr::write_csv(fishery.office.table, here::here("outputs","fishery_offices.csv"))

fisheryoffice <- fishery.office.table
  
usethis::use_data(fisheryoffice, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(fisheryoffice)
cat(data.description, file=here::here("R",paste0("data-","fisheryoffice")))

#rebuild and load the package
devtools::load_all() # restarts and loads




```

Get locations from catch records
```{r}

locs.pesca <- clean_locs_pesca()

data("catchdata")

data("inegigc")

catch.locations.data <- get_locations_catch(catchdata) 

readr::write_csv(catch.locations.data, here::here("outputs","catch_locations.csv"))

catchlocations <- catch.locations.data %>% 
  dplyr::bind_rows() 

usethis::use_data(catchlocations, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(catchlocations)
cat(data.description, file=here::here("R",paste0("data-","catchlocations")))

#rebuild and load the package
devtools::load_all() # restarts and loads


```


#Create table georeferrenced locations from INEGI & Google and add distance to nearest fishery office

```{r location table, include=FALSE}

locs.pesca <- clean_locs_pesca()

correct.ent <- locs.pesca %>%
    select(-NOM_LOC_REV)

correct.loc <- locs.pesca %>%
    select(-NOM_ENT_REV) 

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

#will extract all locations within states, not only coastal areas
#only run once will create attached data
#get_inegi(ents.coast.selecc)

data("inegigc")

locs.census <- inegigc %>% 
   select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)
  
#get georeferenced locations from inegi
locs.inegi <- googlesheets4::read_sheet("1Nww_0aSf2yQv9YqK8LqmJmIM2wwpHZCSpG1NcGfIfHQ", sheet = "busqueda_inegi") %>% 
  dplyr::rename(deci_lat=latitud, deci_lon=longitud) %>% 
  keep_when(utilizar == 1) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
  select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)

#test for duplicates 
dupes.inegi <- locs.inegi %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1) %>% 
  keep_when(dupe == TRUE)

#get georeferenced locations from google
locs.google <- googlesheets4::read_sheet("1Nww_0aSf2yQv9YqK8LqmJmIM2wwpHZCSpG1NcGfIfHQ", sheet = "busqueda_google") %>% 
  dplyr::rename(NOM_LOC=NOM_LOC_REV, deci_lat=lat, deci_lon=lon) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
  keep_when(utilizar == 1) %>% 
  mutate(deci_lon=as.numeric(deci_lon), deci_lat=as.numeric(deci_lat)) %>% 
  select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)

#test for duplicates 
dupes.google <- locs.google %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1)

#Atlas fishing locations Pacific
locs.pesca.pacifico <- googlesheets4::read_sheet("1KxO30FqYed655frJDiLobx3N-slFNpNLpHnbxE4cLcI") %>% 
  dplyr::rename(NOM_ENT= ESTADO, NOM_LOC = LOCALIDAD_REVISADA) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  select(-LOCALIDAD_ORIGINAL, -ID) %>% 
  dplyr::rename(deci_lat =LATITUD, deci_lon=LONGITUD) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
   select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)


#test for duplicates 
dupes.pacifico <- locs.pesca.pacifico %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1)

#combine all locations
georef.locs <- dplyr::bind_rows(locs.census, locs.inegi, locs.google, locs.pesca.pacifico) %>% 
  distinct(NOM_ENT, NOM_LOC, deci_lat, deci_lon) 

#make spatial

georef.locs.sp <- georef.locs %>% 
  sf::st_as_sf(coords = c("deci_lon", "deci_lat"), crs = 4326) %>% 
  sf::st_set_crs(4326)

#Add distance to nearest fishery office to help resolve duplicates when georeferincing a catch 

fishery.offices <- sf::st_read(here::here("outputs","fishery_offices.shp")) %>%   sf::st_transform(crs = 4326) %>% 
  sf::st_centroid() 

num.georef.locs <- 1:nrow(georef.locs.sp)

georef.locs.dist <- parallel::mclapply(num.georef.locs, get_distance, fishery.offices) %>% 
  dplyr::bind_rows() 


georeflocskm <- georef.locs.dist %>% 
  mutate(distance_km = distance / 1000)

readr::write_csv(georeflocskm, here::here("outputs","georef_locations.csv"))

usethis::use_data(georeflocskm, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(georeflocskm)
cat(data.description, file=here::here("R",paste0("data-","georeflocskm")))

#rebuild and load the package
devtools::load_all() # restarts and loads

```


Georeference unique locations
```{r}

data("catchlocations")

data("georeflocskm")


georefcatchlocs <- georef_catch(catchlocations, georeflocskm)
  
usethis::use_data(georefcatchlocs, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(georefcatchlocs)
cat(data.description, file=here::here("R",paste0("data-","georefcatchlocs")))

#rebuild and load the package
devtools::load_all() # restarts and loads

```

Get permit information from the Registro Nacional de Pesca y Acuacultura RNPA and georeference
https://www.gob.mx/conapesca/documentos/registro-nacional-de-pesca-y-acuacultura-rnpa

```{r get catch, message=TRUE, warning=TRUE, include=FALSE}

data("georeflocskm")

locs.pesca <- clean_locs_pesca()

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

inegi.cost.comm.all  <- readr::read_csv(here::here("inputs","all_coastal_communities_INEGI.csv")) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_MUN = stringi::stri_trans_general(str = NOM_MUN, id = "Latin-ASCII"), NOM_MUN = toupper(NOM_MUN)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  keep_when(NOM_ENT %in% coast.selecc)

inegi.cost.mun <- inegi.cost.comm.all %>% 
  distinct(NOM_MUN, CVE_MUN)

permit.files <- list.files(path = here::here("data-raw","permit_data"), full.names = TRUE)
no.permitfiles <- 1:length(permit.files)

permit.data <- lapply(no.permitfiles, extract_permits, permit.files, coast.selecc, locs.pesca, georeflocskm)

permitdata <- permit.data %>% 
  dplyr::bind_rows() %>% 
  distinct(rnp_code, NOM_LOC, NOM_ENT, deci_lat, deci_lon, fishery_office, fishery_office_ent, distance_km)


usethis::use_data(permitdata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(permitdata)
cat(data.description, file=here::here("R",paste0("data-","permitdata")))

#rebuild and load the package
devtools::load_all() # restarts and loads

readr::write_csv(permitdata, here::here("outputs","rnp_fishery_permits.csv"))


```



Georeference all catch data
```{r}

data("georefcatchlocs")
data("catchdata")
data("permitdata")

catch.plot.inter <- georef_catch_data(georefcatchlocs, catchdata)

#plotly::ggplotly(catch.plot.inter)

```
Grid for environmental data
```{r}

catch.geo.pacific <- readr::read_csv(here::here("outputs","georefcatch_loc_pacific.csv"))
 

catch.pacific.spatial <- sf::st_as_sf(catch.geo.pacific, coords = c("deci_lon", "deci_lat"), crs = 4326)

sf::sf_use_s2(FALSE)

gc.polygon <- sf::st_read(here::here("data-raw","SHP","Golfo_california_wetland_poly_WGS84.shp")) %>%   
    sf::st_transform(crs = 4269) %>% 
    sf::st_make_valid()
#make grid at scale of fines oceanographic data

#simplify GOC polygon otherwise it takes too much memory for intersection
#https://www.r-bloggers.com/2021/03/simplifying-geospatial-features-in-r-with-sf-and-rmapshaper/
gc.polygon.simple <- sf::st_simplify(mv, preserveTopology = FALSE, dTolerance = 1000)

#0.083° × 0.083° degree grid Global Ocean Physics Reanalysis
#31 Dec 1992 to 26 Feb 2024 50 vertical levels
#https://data.marine.copernicus.eu/product/GLOBAL_MULTIYEAR_PHY_001_030/description
#0.25° × 0.25° degree grid Global Ocean Biogeochemistry Hindcast
#1993-2019 75 vertical levels

gc.polygon.grid <- sf::st_make_grid(gc.polygon.simple, cellsize = c(0.083, 0.083), what = "polygons") %>% 
  sf::st_as_sf() %>% 
  sf::st_set_crs(4269) 

#CONTINUE HERE
#could also calculate distance to nearest grid point and then el;iminate points that fall in land with a land mass polygon
gc.ocean.grid <- sf::st_intersection(gc.polygon.grid, gc.polygon.simple) 

gc.grid.plot <- ggplot2::ggplot(gc.polygon.grid) + 
    ggplot2::geom_sf() +
    ggplot2::geom_sf(data = gc.polygon, color = "red", alpha=0.4) +
    ggplot2::geom_sf(data = catch.pacific.spatial, color = "darkblue") +
    ggplot2::labs(x = "Longitude",
                  y="Latitude",
                  title = "Georeferenced catch data",
                  subtitle = "Grid 0.083° × 0.083° degree")

gc.grid.plot

get_pointdistance<- function(eachcatchpoint, catch.pacific.spatial, gc.ocean.grid){
  
  this.catch.spatial <- catch.pacific.spatial[eachcatchpoint,]
  
  this.distance <- st_distance(this.catch.spatial, gc.ocean.grid)
  
}


  # 
```

