---
title: "Landing sites"
output: html_notebook
---


```{r load package, include=FALSE}
#rebuild and load the package
devtools::load_all() # restarts and loads

# List of packages for session
.packages = c("import")

# Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])

# Load packages into session 
lapply(.packages, require, character.only=TRUE)

#use this for packages
#specifies what functions to import
# https://cran.r-project.org/web/packages/import/vignettes/import.html
import::from(dplyr, mutate, distinct, select, group_by, summarise, keep_when = filter, rename)

#to fix access with google
#https://github.com/orgs/community/discussions/39210

```

Get catch from official record. 
Filter out catch for artisanal fisheries and eliminate aquaculture production.
Uses table with corrected names to standarize names of locations and fishery offices and correct incorrect state names.
Combines catch files for all years and selects landing site, state, fishery office, species, weight, and catch value.
Resulting data is imported as catch rmd into the package

Need to add separation of sites that are closer on the Pacific Coast that the Gulf after the locations are georreferenced

```{r get catch data, include=FALSE}

locs.pesca <- clean_locs_pesca()

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

#Currently only using data 2018-2023, other data have too many inconsistencies
#catch.files <- list.files(path = here::here("data-raw","original_fisheries_data"), full.names = TRUE)
catch.files <- list.files(path = here::here("data-raw","original_fisheries_data"), full.names = TRUE, pattern = "2018|2019|2020|2021|2022|2023")

no.catchfiles <- 1:length(catch.files)

#NOTE: There are 16,531 records between 2000-2004 with 0 price for catch

#parallel::mclapply(no.catchfiles, get_catch, catch.files, coast.selecc, locs.pesca)
#use lapply for debugging
lapply(no.catchfiles, get_catch, catch.files, coast.selecc, locs.pesca)

catch.files.clean <- list.files(path = here::here("data-raw","fisheries_data"), full.names = TRUE)
no.catchfiles.clean <- 1:length(catch.files.clean)

catchdata <- lapply(no.catchfiles.clean, clean_catch, catch.files.clean) %>% 
  dplyr::bind_rows() %>% 
  mutate(catch_id = paste0("catch_",1:nrow(.)))

usethis::use_data(catchdata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(catchdata)
cat(data.description, file=here::here("R",paste0("data-","catchdata")))

#rebuild and load the package
devtools::load_all() # restarts and loads


```


Gets list of fishery offices from catch records
Uses standarized catch data to get unique fishery offices and georreference them
```{r}

locs.pesca <- clean_locs_pesca()

#get unique georreferenced rnp codes

data("inegigc")
data("catchdata")



fishery.offices.list <- get_fishery_office_catch(catchdata) 

#fishery office list for 2024 is available here https://www.conapesca.gob.mx/work/sites/cona/dgppe/Subdelegaciones_de_Pesca.pdf
#the La Reforma office is in the Municipality of Angostura and has CVE_LOC = 250020093
#there are locations named La Reforma sites in Sinaloa so had to be extracted manually

fishery.office.table <- fishery.offices.list %>% 
  dplyr::bind_rows() %>% 
  dplyr::arrange(NOM_LOC,NOM_ENT)

fishery.spatial <- sf::st_as_sf(fishery.office.table, coords = c("deci_lon", "deci_lat"), crs = 4326) 

fishery.spatial %>% 
  sf::st_write(here::here("outputs","fishery_offices.shp"), append=FALSE)

readr::write_csv(fishery.office.table, here::here("outputs","fishery_offices.csv"))

fisheryoffice <- fishery.office.table
  
usethis::use_data(fisheryoffice, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(fisheryoffice)
cat(data.description, file=here::here("R",paste0("data-","fisheryoffice")))

#rebuild and load the package
devtools::load_all() # restarts and loads




```

Get locations from catch records
```{r}

locs.pesca <- clean_locs_pesca()

data("catchdata")

data("inegigc")

catch.locations.data <- get_locations_catch(catchdata) 

readr::write_csv(catch.locations.data, here::here("outputs","catch_locations.csv"))

catchlocations <- catch.locations.data %>% 
  dplyr::bind_rows() 

usethis::use_data(catchlocations, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(catchlocations)
cat(data.description, file=here::here("R",paste0("data-","catchlocations")))

#rebuild and load the package
devtools::load_all() # restarts and loads


```


#Create table georeferrenced locations from INEGI & Google and add distance to nearest fishery office

```{r location table, include=FALSE}

locs.pesca <- clean_locs_pesca()

correct.ent <- locs.pesca %>%
    select(-NOM_LOC_REV)

correct.loc <- locs.pesca %>%
    select(-NOM_ENT_REV) 

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

#will extract all locations within states, not only coastal areas
#only run once will create attached data
#get_inegi(ents.coast.selecc)

data("inegigc")

locs.census <- inegigc %>% 
   select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)
  
#get georreferenced locations from inegi
locs.inegi <- googlesheets4::read_sheet("1Nww_0aSf2yQv9YqK8LqmJmIM2wwpHZCSpG1NcGfIfHQ", sheet = "busqueda_inegi") %>% 
  dplyr::rename(deci_lat=latitud, deci_lon=longitud) %>% 
  keep_when(utilizar == 1) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
  select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)

#test for duplicates 
dupes.inegi <- locs.inegi %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1) %>% 
  keep_when(dupe == TRUE)

#get georreferenced locations from google
locs.google <- googlesheets4::read_sheet("1Nww_0aSf2yQv9YqK8LqmJmIM2wwpHZCSpG1NcGfIfHQ", sheet = "busqueda_google") %>% 
  dplyr::rename(NOM_LOC=NOM_LOC_REV, deci_lat=lat, deci_lon=lon) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
  keep_when(utilizar == 1) %>% 
  mutate(deci_lon=as.numeric(deci_lon), deci_lat=as.numeric(deci_lat)) %>% 
  select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)

#test for duplicates 
dupes.google <- locs.google %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1)

#Atlas fishing locations Pacific
locs.pesca.pacifico <- googlesheets4::read_sheet("1KxO30FqYed655frJDiLobx3N-slFNpNLpHnbxE4cLcI") %>% 
  dplyr::rename(NOM_ENT= ESTADO, NOM_LOC = LOCALIDAD_REVISADA) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  select(-LOCALIDAD_ORIGINAL, -ID) %>% 
  dplyr::rename(deci_lat =LATITUD, deci_lon=LONGITUD) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
   select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)


#test for duplicates 
dupes.pacifico <- locs.pesca.pacifico %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1)

#combine all locations
georef.locs <- dplyr::bind_rows(locs.census, locs.inegi, locs.google, locs.pesca.pacifico) %>% 
  distinct(NOM_ENT, NOM_LOC, deci_lat, deci_lon) 

#make spatial

georef.locs.sp <- georef.locs %>% 
  sf::st_as_sf(coords = c("deci_lon", "deci_lat"), crs = 4326) %>% 
  sf::st_set_crs(4326)

#Add distance to nearest fishery office to help resolve duplicates when georreferincing a catch 

fishery.offices <- sf::st_read(here::here("outputs","fishery_offices.shp")) %>%   sf::st_transform(crs = 4326) %>% 
  sf::st_centroid() 

num.georef.locs <- 1:nrow(georef.locs.sp)

georef.locs.dist <- parallel::mclapply(num.georef.locs, get_distance, fishery.offices) %>% 
  dplyr::bind_rows() 


georeflocskm <- georef.locs.dist %>% 
  mutate(distance_km = distance / 1000)

readr::write_csv(georeflocskm, here::here("outputs","georef_locations.csv"))

usethis::use_data(georeflocskm, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(georeflocskm)
cat(data.description, file=here::here("R",paste0("data-","georeflocskm")))

#rebuild and load the package
devtools::load_all() # restarts and loads

```


Georreference unique locations
```{r}

data("catchlocations")

data("georeflocskm")


georeffcatchlocs <- georref_catch(catchlocations, georeflocskm)
  
usethis::use_data(georeffcatchlocs, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(georeffcatchlocs)
cat(data.description, file=here::here("R",paste0("data-","georeffcatchlocs")))

#rebuild and load the package
devtools::load_all() # restarts and loads

```

Georeference all catch data
```{r}

data("georeffcatchlocs")

data("catchdata")

#buffer Pacific to eliminate communities not in Gulf of California
pac.buffer <- sf::st_read(here::here("data-raw","pac_mx_gc.shp")) %>%   
  sf::st_transform(crs = 4326) %>% 
  sf::st_buffer(0.1)

plot(pac.buffer)

georeff_catch_data()
```


Get permit information from the Registro Nacional de Pesca y Acuacultura RNPA and georreference
https://www.gob.mx/conapesca/documentos/registro-nacional-de-pesca-y-acuacultura-rnpa

```{r get catch, message=TRUE, warning=TRUE, include=FALSE}

georef.locs.km <- readr::read_csv(here::here("outputs","georef_locations.csv")) 


#googlesheet to keep track of names that need correction
#vulnerability_tables
    
googlesheets4::gs4_deauth()

 locs.pesca <- googlesheets4::read_sheet("1eJXjwlHqEzSAZaTmE_XyN93fu6EI0I8igRH4AdvXOl8") %>% 
  mutate(NOM_LOC= stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_LOC_REV= stringi::stri_trans_general(str = NOM_LOC_REV, id = "Latin-ASCII"), NOM_LOC_REV = toupper(NOM_LOC_REV)) %>% 
  mutate(NOM_ENT_REV= stringi::stri_trans_general(str = NOM_ENT_REV, id = "Latin-ASCII"), NOM_ENT_REV = toupper(NOM_ENT_REV)) %>%
   distinct(NOM_LOC, NOM_ENT, NOM_LOC_REV, NOM_ENT_REV) %>% 
   keep_when(!is.na(NOM_LOC_REV))


ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

inegi.cost.comm.all  <- readr::read_csv(paste0("~/coastalvulnerability-inputs/processed_data/all_coastal_communities_INEGI.csv")) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_MUN = stringi::stri_trans_general(str = NOM_MUN, id = "Latin-ASCII"), NOM_MUN = toupper(NOM_MUN)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  keep_when(NOM_ENT %in% coast.selecc)

inegi.cost.mun <- inegi.cost.comm.all %>% 
  distinct(NOM_MUN, CVE_MUN)

permit.files <- list.files(path = here::here("data-raw","permit_data"), full.names = TRUE)
no.permitfiles <- 1:length(permit.files)

permit.data <- lapply(no.permitfiles, extract_permits, permit.files, inegi.cost.comm.all, coast.selecc, locs.pesca, georef.locs.km)

permit.data.rev <- permit.data %>% 
  dplyr::bind_rows() %>% 
  distinct(rnp_code, NOM_LOC, NOM_ENT, deci_lat, deci_lon, fishery_office, fishery_office_ent, distance_km)


readr::write_csv(permit.data.rev, here::here("outputs","rnp_fishery_permits.csv"))


```

Get RNP codes from catch records and georreference
```{r}

georef.locs.km <- readr::read_csv(here::here("outputs","georef_locations.csv"))

googlesheets4::gs4_deauth()

 locs.pesca <- googlesheets4::read_sheet("1eJXjwlHqEzSAZaTmE_XyN93fu6EI0I8igRH4AdvXOl8") %>% 
  mutate(NOM_LOC= stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_LOC_REV= stringi::stri_trans_general(str = NOM_LOC_REV, id = "Latin-ASCII"), NOM_LOC_REV = toupper(NOM_LOC_REV)) %>% 
  mutate(NOM_ENT_REV= stringi::stri_trans_general(str = NOM_ENT_REV, id = "Latin-ASCII"), NOM_ENT_REV = toupper(NOM_ENT_REV)) %>%
   distinct(NOM_LOC, NOM_ENT, NOM_LOC_REV, NOM_ENT_REV) %>% 
   keep_when(!is.na(NOM_LOC_REV))
 
#georeferrenced rnp from lists of fishery permit 
permit.locs.geo <- readr::read_csv(here::here("outputs","rnp_fishery_permits.csv")) 

fishery.office.table <- readr::read_csv(here::here("outputs","fishery_offices.csv")) %>% 
  rename(fishery_office_ent = NOM_ENT, fishery_office = NOM_LOC) %>%
  select(fishery_office_ent, fishery_office, deci_lat, deci_lon)

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

catch.files <- list.files(path = here::here("data-raw","fisheries_data"), full.names = TRUE)
no.catchfiles <- 1:length(catch.files)

#get unique georreferenced rnp codes

catch.rnpa <- lapply(no.catchfiles, get_rnpa_catch, catch.files, coast.selecc, locs.pesca, permit.locs.geo, georef.locs.km, fishery.office.table) 

catch.rnpa.codes <- catch.rnpa %>%
  dplyr::bind_rows() %>% 
  distinct(rnp_code, NOM_ENT, NOM_LOC, deci_lat, deci_lon)
   distinct(rnp_code, NOM_LOC, NOM_ENT, deci_lat, deci_lon, fishery_office, fishery_office_ent, distance_km)

catch.rnpa.codes %>% dplyr::arrange(NOM_ENT, NOM_LOC) %>% View()



```


Georreference landing records
```{r}

permit.locs.geo <- readr::read_csv(here::here("outputs","rnp_fishery_permits.csv"))

georef.locs.all <- readr::read_csv(here::here("outputs","georef_locations.csv")) %>% 
  dplyr::rename(dec_lat_google = deci_lat, dec_lon_google = deci_lon)


coast.selecc <- "SONORA"

catch.files <- list.files(path = here::here("data-raw","fisheries_data"), full.names = TRUE)
no.catchfiles <- 1:length(catch.files)

#match to rnpa number

catch.rnpa <- lapply(no.catchfiles,get_rnpa_catch, catch.files, coast.selecc, permit.locs.geo, georef.locs.all) 

rnpa.catch.table <- catch.rnpa %>%
  dplyr::bind_rows() %>% 
  mutate(rnp_code = as.character(rnp_code))

catch.rnpa.codes <- rnpa.catch.table %>%
  distinct(rnp_code) 

catch.permit.table <- catch.rnpa.codes %>% 
  dplyr::left_join(permit.table, by = c("rnp_code"))

catch.loc <- catch.permit.table %>% 
  dplyr::left_join(rnpa.catch.table, by = c("rnp_code")) %>% 
  select(-NOM_MUN) %>% 
  dplyr::left_join(inegi.cost.mun, by = "CVE_MUN") %>% 
  group_by(CVE_MUN, CVE_LOC, POBTOT) %>% 
  summarise(tot_catch_mt = sum(tot_weight_kg)/ 1000) %>% 
  mutate(catch_percapita = tot_catch_mt/ POBTOT) 

readr::write_csv(catch.loc, paste0(input.dir,"/processed_data/sensitivity/fishery_catch.csv"))


```




Assign catch to locations

```{r location catch, eval=FALSE, include=FALSE}

#googlesheet to keep track of names that need correction
googlesheets4::gs4_deauth()
locs.pesca <- googlesheets4::read_sheet("1eJXjwlHqEzSAZaTmE_XyN93fu6EI0I8igRH4AdvXOl8")

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco", "Yucatán", "Campeche", "Quintana Roo"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")
coast.selecc.pac <- c("BAJA CALIFORNIA SUR","BAJA CALIFORNIA","SONORA","SINALOA","NAYARIT","JALISCO")


#all coastal communities, including those with no census data
inegi.cost.comm.all  <- readr::read_csv(paste0(input.dir,"/processed_data/","all_coastal_communities_INEGI.csv")) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_MUN = stringi::stri_trans_general(str = NOM_MUN, id = "Latin-ASCII"), NOM_MUN = toupper(NOM_MUN)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC))

catch.files <- list.files(path = paste0(input.dir,"/catch"), full.names = TRUE)
no.catchfiles <- 1:length(catch.files)

catch.locations <- lapply(no.catchfiles, extract_locations, catch.files, coast.selecc, locs.pesca, inegi.cost.comm.all)
  
catch.locs.all <- catch.locations %>% 
  dplyr::bind_rows() %>% 
  distinct(NOM_ENT, NOM_LOC)

#check for duplicates
# catch.locs.inegi.all <-  catch.locs.all %>% 
#   group_by(NOM_ENT, NOM_LOC) %>%
#   mutate(dupe = dplyr::n()>1)
  
catch.locs.all %>% 
    readr::write_csv(paste0(input.dir,"/processed_data/localidades_captura_inegi.csv"))
 
  #subset locations in the pacific and match with landing sites
  catch.locs.pac <- catch.locs.all %>% 
    keep_when(NOM_ENT %in% coast.selecc.pac) %>% 
    dplyr::left_join(loc.pacifico, by = c("NOM_ENT","NOM_LOC"))
     
  #used to find duplicates https://stackoverflow.com/questions/6986657/find-duplicated-rows-based-on-2-columns-in-data-frame-in-r
  #there are multiple landing codes by location, so not using
  # catch.file.sites.table %>% 
  #   group_by(NOM_ENT, NOM_LOC) %>%
  #   mutate(dupe = dplyr::n()>1)
  # 
  
  
   
  
```

