---
title: "Georeference landing sites and plot catch for COBI 2026 FIP project"
output: html_notebook
---


```{r load package, include=FALSE}
#rebuild and load the package
devtools::load_all() # restarts and loads

# List of packages for session
.packages = c("import")

# Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])

# Load packages into session 
lapply(.packages, require, character.only=TRUE)

#use this for packages
#specifies what functions to import
# https://cran.r-project.org/web/packages/import/vignettes/import.html
import::from(dplyr, mutate, distinct, select, group_by, summarise, keep_when = filter, rename)

#to fix access with google
#https://github.com/orgs/community/discussions/39210

#to fix access issues
#https://www.namehero.com/blog/solve-temporary-failure-in-name-resolution-error-in-linux/


```

Get catch from official record. 
Filter out catch for artisanal fisheries and eliminate aquaculture production.
Uses table with corrected names to standarize names of locations and fishery offices and correct incorrect state names.
Combines catch files for all years and selects landing site, state, fishery office, species, weight, and catch value.
Resulting data is imported as catch rmd into the package

```{r get catch data, echo=TRUE}

locs.pesca <- clean_locs_pesca()

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

#Currently only using data 2018-2023, other data have too many inconsistencies
#catch.files <- list.files(path = here::here("data-raw","original_fisheries_data"), full.names = TRUE)
catch.files <- list.files(path = here::here("data-raw","original_fisheries_data"), full.names = TRUE, pattern = "2018|2019|2020|2021|2022|2023|2024|2025")

no.catchfiles <- 1:length(catch.files)

#NOTE: There are 16,531 records between 2000-2004 with 0 price for catch

#parallel::mclapply(no.catchfiles, get_catch, catch.files, coast.selecc, locs.pesca)
#use lapply for debugging
lapply(no.catchfiles, get_catch, catch.files, coast.selecc, locs.pesca)

catch.files.clean <- list.files(path = here::here("data-raw","fisheries_data"), full.names = TRUE)
no.catchfiles.clean <- 1:length(catch.files.clean)

catchdata <- lapply(no.catchfiles.clean, clean_catch, catch.files.clean) %>% 
  dplyr::bind_rows() %>% 
  mutate(catch_id = paste0("catch_",1:nrow(.)))

usethis::use_data(catchdata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(catchdata)
cat(data.description, file=here::here("R",paste0("data-","catchdata")))

#rebuild and load the package
devtools::load_all() # restarts and loads


```


Gets list of fishery offices from catch records
Uses standarized catch data to get unique fishery offices and georeference them
```{r Get fishery offices}

locs.pesca <- clean_locs_pesca()

#get unique georeferenced rnp codes

data("inegigc")
data("catchdata")

fishery.offices.list <- get_fishery_office_catch(catchdata) 

#fishery office list for 2024 is available here https://www.conapesca.gob.mx/work/sites/cona/dgppe/Subdelegaciones_de_Pesca.pdf
#the La Reforma office is in the Municipality of Angostura and has CVE_LOC = 250020093
#there are locations named La Reforma sites in Sinaloa so had to be extracted manually

fishery.office.table <- fishery.offices.list %>% 
 dplyr::left_join(inegigc) %>% 
  dplyr::select(-POBTOT, -POBFEM,  -POBMAS)

fishery.spatial <- sf::st_as_sf(fishery.office.table, coords = c("deci_lon", "deci_lat"), crs = 4326) %>% 
  sf::st_transform(4269)

fishery.spatial %>% 
  sf::st_write(here::here("outputs","fishery_offices.shp"), append=FALSE)

readr::write_csv(fishery.office.table, here::here("outputs","fishery_offices.csv"))

world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")

point_extent <- sf::st_bbox(fishery.spatial)

ggplot2::ggplot() +
   ggplot2::geom_sf(data = world, fill = "antiquewhite", color = "gray") + # Base layer
  ggplot2::geom_sf(data = fishery.spatial, size = 2, color = "blue") +
  ggplot2::coord_sf(xlim = c(point_extent["xmin"], point_extent["xmax"]), 
           ylim = c(point_extent["ymin"], point_extent["ymax"]),
           expand = TRUE) +   
  ggplot2::theme_minimal()

fisheryoffice <- fishery.office.table
  
usethis::use_data(fisheryoffice, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(fisheryoffice)
cat(data.description, file=here::here("R",paste0("data-","fisheryoffice")))

#rebuild and load the package
devtools::load_all() # restarts and loads




```

Get locations from catch records
```{r Get locations from catch, echo=TRUE}

locs.pesca <- clean_locs_pesca()

data("catchdata")

data("inegigc")

catch.locations.data <- get_locations_catch(catchdata) 

readr::write_csv(catch.locations.data, here::here("outputs","catch_locations.csv"))

catchlocations <- catch.locations.data %>% 
  dplyr::bind_rows() 

usethis::use_data(catchlocations, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(catchlocations)
cat(data.description, file=here::here("R",paste0("data-","catchlocations")))

#rebuild and load the package
devtools::load_all() # restarts and loads


```


#Create table georeferrenced locations from INEGI & Google and add distance to nearest fishery office

```{r location table, echo=TRUE}

locs.pesca <- clean_locs_pesca()

correct.ent <- locs.pesca %>%
    select(-NOM_LOC_REV)

correct.loc <- locs.pesca %>%
    select(-NOM_ENT_REV) 

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

#will extract all locations within states, not only coastal areas
#only run once will create attached data
#get_inegi(ents.coast.selecc)

data("inegigc")

locs.census <- inegigc %>% 
   select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)
  
#get georeferenced locations from inegi
locs.inegi <- googlesheets4::read_sheet("1Nww_0aSf2yQv9YqK8LqmJmIM2wwpHZCSpG1NcGfIfHQ", sheet = "busqueda_inegi") %>% 
  dplyr::rename(deci_lat=latitud, deci_lon=longitud) %>% 
  keep_when(utilizar == 1) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
  select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)

#test for duplicates 
dupes.inegi <- locs.inegi %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1) %>% 
  keep_when(dupe == TRUE)

#get georeferenced locations from google
locs.google <- googlesheets4::read_sheet("1Nww_0aSf2yQv9YqK8LqmJmIM2wwpHZCSpG1NcGfIfHQ", sheet = "busqueda_google") %>% 
  dplyr::rename(NOM_LOC=NOM_LOC_REV, deci_lat=lat, deci_lon=lon) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
  keep_when(utilizar == 1) %>% 
  mutate(deci_lon=as.numeric(deci_lon), deci_lat=as.numeric(deci_lat)) %>% 
  select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)

#test for duplicates 
dupes.google <- locs.google %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1)

#Atlas fishing locations Pacific
locs.pesca.pacifico <- googlesheets4::read_sheet("1KxO30FqYed655frJDiLobx3N-slFNpNLpHnbxE4cLcI") %>% 
  dplyr::rename(NOM_ENT= ESTADO, NOM_LOC = LOCALIDAD_REVISADA) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  select(-LOCALIDAD_ORIGINAL, -ID) %>% 
  dplyr::rename(deci_lat =LATITUD, deci_lon=LONGITUD) %>% 
  keep_when(NOM_ENT %in% coast.selecc) %>% 
   select(NOM_ENT, NOM_LOC, deci_lat, deci_lon)


#test for duplicates 
dupes.pacifico <- locs.pesca.pacifico %>% 
  group_by(NOM_ENT, NOM_LOC) %>% 
  mutate(dupe = dplyr::n()>1)

#combine all locations
georef.locs <- dplyr::bind_rows(locs.census, locs.inegi, locs.google, locs.pesca.pacifico) %>% 
  distinct(NOM_ENT, NOM_LOC, deci_lat, deci_lon) 

#make spatial

georef.locs.sp <- georef.locs %>% 
  sf::st_as_sf(coords = c("deci_lon", "deci_lat"), crs = 4326) %>% 
  sf::st_transform(4269)

#Add distance to nearest fishery office to help resolve duplicates when georeferincing a catch 

fishery.offices <- sf::st_read(here::here("outputs","fishery_offices.shp")) 

num.georef.locs <- 1:nrow(georef.locs.sp)

#georef.locs.dist <- parallel::mclapply(num.georef.locs, get_distance, fishery.offices) 

georef.locs.dist <- lapply(num.georef.locs, get_distance, fishery.offices) 

georeflocskm <- georef.locs.dist %>% 
 dplyr::bind_rows()

readr::write_csv(georeflocskm, here::here("outputs","georef_locations.csv"))

usethis::use_data(georeflocskm, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(georeflocskm)
cat(data.description, file=here::here("R",paste0("data-","georeflocskm")))

#rebuild and load the package
devtools::load_all() # restarts and loads

```


Georeference unique locations
```{r Georeference locations, echo=TRUE}

data("catchlocations")

data("georeflocskm")


georefcatchlocs <- georef_catch(catchlocations, georeflocskm)
  
usethis::use_data(georefcatchlocs, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(georefcatchlocs)
cat(data.description, file=here::here("R",paste0("data-","georefcatchlocs")))

#rebuild and load the package
devtools::load_all() # restarts and loads

```

Get permit information from the Registro Nacional de Pesca y Acuacultura RNPA and georeference
https://www.gob.mx/conapesca/documentos/registro-nacional-de-pesca-y-acuacultura-rnpa

```{r Get permit data, echo=TRUE, message=TRUE, warning=TRUE}

data("georeflocskm")

locs.pesca <- clean_locs_pesca()

ents.coast.selecc <- toupper(c("Baja California Sur", "Baja California", "Sonora", "Sinaloa", "Nayarit", "Jalisco"))
coast.selecc <- stringi::stri_trans_general(str = ents.coast.selecc, id = "Latin-ASCII")

inegi.cost.comm.all  <- readr::read_csv(here::here("inputs","all_coastal_communities_INEGI.csv")) %>% 
  mutate(NOM_ENT= stringi::stri_trans_general(str = NOM_ENT, id = "Latin-ASCII"), NOM_ENT = toupper(NOM_ENT)) %>% 
  mutate(NOM_MUN = stringi::stri_trans_general(str = NOM_MUN, id = "Latin-ASCII"), NOM_MUN = toupper(NOM_MUN)) %>% 
  mutate(NOM_LOC = stringi::stri_trans_general(str = NOM_LOC, id = "Latin-ASCII"), NOM_LOC = toupper(NOM_LOC)) %>% 
  keep_when(NOM_ENT %in% coast.selecc)

inegi.cost.mun <- inegi.cost.comm.all %>% 
  distinct(NOM_MUN, CVE_MUN)

permit.files <- list.files(path = here::here("data-raw","permit_data"), full.names = TRUE)
no.permitfiles <- 1:length(permit.files)

permit.data <- lapply(no.permitfiles, extract_permits, permit.files, coast.selecc, locs.pesca, georeflocskm)

permitdata <- permit.data %>% 
  dplyr::bind_rows() %>% 
  distinct(rnp_code, NOM_LOC, NOM_ENT, deci_lat, deci_lon, fishery_office, fishery_office_ent, distance_km)


usethis::use_data(permitdata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(permitdata)
cat(data.description, file=here::here("R",paste0("data-","permitdata")))

#rebuild and load the package
devtools::load_all() # restarts and loads

readr::write_csv(permitdata, here::here("outputs","rnp_fishery_permits.csv"))


```


Georeference all catch data
```{r Georeference catch, echo=TRUE}

data("georefcatchlocs")
data("catchdata")
data("permitdata")

if(file.exists("outputs/gpkg_grid.gpkg")) file.remove("outputs/gpkg_grid.gpkg")

catch.plot.inter <- georef_catch_data(georefcatchlocs, catchdata)

#plotly::ggplotly(catch.plot.inter)

catch.plot.inter

```


```{r extractcatchspatial, echo=TRUE}


catch.pacific <- sf::st_read(here::here("outputs","gpkg_grid.gpkg"), layer='goc_catch') %>% 
    sf::st_transform(crs = 6362)
 
rgi.shape <- sf::st_read(here::here("data-raw","SHP","Región de las Grandes Islas (1).shp"))

rgi.shape.buffer <- rgi.shape %>% 
    sf::st_transform(crs = 6362) %>% 
  sf::st_buffer(75000)

sf::st_write(rgi.shape.buffer, here::here("outputs","rgi_buffer.shp"), delete_layer=TRUE)

 plot.buffer <- ggplot2::ggplot(rgi.shape.buffer) + 
    ggplot2::geom_sf(fill = "lightgreen") +
    ggplot2::geom_sf(data = rgi.shape, color = "darkred") +
    ggplot2::geom_sf(data = catch.pacific, color = "darkblue") +
      ggplot2::labs(x = "Longitude",
                  y="Latitude") +
   ggspatial::annotation_north_arrow(
    location = "bl",             # "bl" = bottom left, "tr" = top right, etc.
    which_north = "true",        # "true" (North Pole) or "grid" (top of map)
    pad_x = ggplot2::unit(0.2, "in"),     # Horizontal distance from edge
    pad_y = ggplot2::unit(0.2, "in"),     # Vertical distance from edge
    style = ggspatial::north_arrow_fancy_orienteering() # Choose arrow visual style
  ) +
   ggthemes::theme_map()
 
 #extract catch points with buffer
 
buffer.catch <- sf::st_intersection(catch.pacific, rgi.shape.buffer)

 sf::st_write(buffer.catch, dsn=here::here("outputs", "gpkg_grid.gpkg"), layer='rgi_catch', layer_options = "OVERWRITE=YES", append = FALSE)

 
 rgi.catch <- sf::st_drop_geometry(buffer.catch) %>% 
   dplyr::select(rnp_code, NOM_LOC, NOM_ENT, fishery_office, species, landed_w_kg, value_mxn, year, month) %>% 
   tidyr::as_tibble()
 
rgi.catch.plot <- rgi.catch %>% 
  dplyr::summarise(tot_w_kg = sum(landed_w_kg), .by=c("species","year","NOM_ENT")) %>% 
  dplyr::mutate(tot_w_ton = tot_w_kg/1000) %>% 
 ggplot2::ggplot(ggplot2::aes(x = year, y = tot_w_ton, fill = NOM_ENT)) +
  ggplot2::geom_col() +
   ggplot2::facet_wrap(~species, scales = "free_y") +
  ggplot2::labs(title = "Captura Región de Grandes Islas", x = "Año", y = "Captura toneladas", fill = "Entidad") 

 rgi.catch %>% data.table::fwrite(here::here("outputs","captura_rgi_2018-2025.csv"))

 #rgi.catch <- data.table::fread(here::here("outputs","captura_rgi_2018-2025.csv"))

```

